Reasoning about a Capability Machine with Local Capabilities

### 2017-06-14 popl-2018/paper.tex
Expand the discussion of local capabilities in the introductory section of the LR.

\paragraph{Local capabilities}
A final thing we explain here is how our results take into account local
capabilities and their special treatment by the hardware. Our idea here is to
introduce a distinction between temporary and permanent invariants. At any point
in a proof, we may decide to revoke temporary invariants. Global capabilities
are then guaranteed to remain safe, but local ones aren't. Concretely, we will
perform such revocations when we need to break the validity of capabilities that
we previously provided to the adversary as a local capability (for example,
return or stack pointers from a previous invocation). In such situations, the
hardware guarantees that global capabilities (which cannot have write-local
permission because of a central assumption\dominique{highlight earlier}) cannot
have been used to store local capabilities and therefore remain safe even after
revoking temporary invariants. Formally, this leads to two distinct future world
relations: we say that $W' \futurewk W$ if a future world $W'$ preserves
temporary invariants and $W' \futurestr W$ if it doesn't. Interestingly, this
is a variant of a notion of public/private future worlds that has been previously
used in the literature (see Section~\ref{sec:discussion}).
%

@@

\noindent \textbf{Reasoning about capability machine programs} \emph{Semantic,
  but not syntactic properties} The logical relation defined in
Section~\ref{sec:logical-relation} allows us to reason about capability machine
programs. A limitation w.r.t. previous work is that the logical relation is
tailored exclusively towards semantic properties, not syntactic ones.

@@

For our purposes, this restriction is unproblematic, since we are only
interested in proving semantic properties (e.g., an assertion will never fail).
However, in other situations, we may be interested in proving more syntactic
properties like the more syntactic properties, that are often considered in
object capability literature like confinement, no authority amplification etc.
Although such properties are typically more restrictive and less easy to use for
reasoning, \citet{Devriese:2016ObjCap} have demonstrated how a logical relation
much like ours can be adapted to also support them, by quantifying the logical
relation over a custom interpretation of effectful computations and the type of
references. We expect their solution can be readily adapted to our setting,
modulo some details (like the fact that we do not just have read-write
capabilities, but also other types).
% \emph{No program logic} TODO: discuss this?

>>>>

\paragraph{Local capabilities}
A final thing we explain here is how our results take into account local
capabilities and their special treatment by the hardware. Normally, when we
invoke an untrusted piece of code and provide it with certain global
capabilities, it may have stored those capabilities in memory and we will only
be able to reinvoke the code if we can guarantee that those values are still
valid. Formally, worlds will represent the invariants that the global
capabilities' safety relies on and reinvocation will only be possible in future
worlds.

However, if we provide the adversary with local capabilities in that first
invocation, then the situation is a bit different. Since the adversary has no
way to store local capabilities, we can allow the second invocation to happen in
\emph{private} future worlds ($W' \futurestr W$), in which global capabilities
remain safe, but local ones don't. This private future world relation is more
liberal than the public one ($W' \futurewk W$, in which \emph{all} safe
capabilities remain safe). In particular, worlds may contain \emph{temporary}
regions, representing invariants that only local capabilties may rely on for
their safety, and which may be revoked (disabled) in private future worlds.

Interestingly, this idea is a variant of a notion of public/private future
worlds that has been previously used in the literature (see
Section~\ref{sec:discussion}). However, temporary regions are new in our setting
and there is an interesting interplay with the recursiveness of the worlds: for
a temporary region, the predicate $H \in \Worlds \rightarrow
\powerset{\MemSegments}$ (which defines which memory segments are safe in the
current state of the world) is only required to be monotonous w.r.t.\ public
future worlds (i.e.\ safe memory segments remain safe in public future worlds),
but for permanent regions, it must be monotonous w.r.t.\ private future worlds.
As a consequence, the memory for a permanent region may not contain local
capabilities (as their safety would be broken in private future worlds), which
in turn implies that only local capabilities may have store-local permission (a
general sanity requirement when using local capabilities\footnote{In fact, local
  capabilities become useless as soon as the adversary has access to a single
  global, store-local capability.}).

@@

\noindent \textbf{Reasoning about capability machine programs}\\
\emph{Semantic, but not syntactic properties} The logical relation defined in
Section~\ref{sec:logical-relation} allows us to reason about capability machine
programs. A limitation w.r.t. previous work is that the logical relation is
tailored exclusively towards semantic properties, not syntactic ones.

@@

For our purposes, this restriction is unproblematic, since we are only
interested in proving semantic properties (e.g., an assertion will never fail).
However, in other situations, we may be interested in proving more syntactic
properties like the more syntactic properties, that are often considered in
object capability literature like confinement, no authority amplification etc.
Although such properties are typically more restrictive and less easy to use for
reasoning, \citet{Devriese:2016ObjCap} have demonstrated how a logical relation
much like ours can be adapted to also support them, by quantifying the logical
relation over a custom interpretation of effectful computations and the type of
references. We expect their solution can be readily adapted to our setting,
modulo some details (like the fact that we do not just have read-write
capabilities, but also other types).\\\\
% \emph{No program logic} TODO: discuss this?
